\documentclass[11pt,letterpaper]{article}

\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{parskip}

% Code listing style
\definecolor{codebg}{rgb}{0.95,0.95,0.95}
\definecolor{codestr}{rgb}{0.13,0.55,0.13}
\definecolor{codecom}{rgb}{0.5,0.5,0.5}
\definecolor{codekw}{rgb}{0.0,0.0,0.8}

\lstset{
  backgroundcolor=\color{codebg},
  basicstyle=\ttfamily\small,
  breaklines=true,
  captionpos=b,
  commentstyle=\color{codecom},
  frame=single,
  keywordstyle=\color{codekw}\bfseries,
  language=bash,
  showstringspaces=false,
  stringstyle=\color{codestr},
  xleftmargin=6pt,
  xrightmargin=6pt,
}

\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  urlcolor=blue,
}

\title{\textbf{pdfuse v0.2.0 — Project Report}\\
  \large New Features: Compress, Rotate, Watermark, Reorder}
\author{Kangshuo Li}
\date{April 2026}

\begin{document}

\maketitle
\tableofcontents
\newpage

% -------------------------------------------------------------------
\section{Overview}
% -------------------------------------------------------------------

\textbf{pdfuse} is a command-line PDF utility written in Python.
Version~0.1.0 (Project~1) provided four core operations:
\texttt{merge}, \texttt{split}, \texttt{convert}, and \texttt{info}.
Version~0.2.0 (this project) adds four new operations:
\texttt{compress}, \texttt{rotate}, \texttt{watermark}, and
\texttt{reorder}.

All commands share the same design philosophy as Project~1:

\begin{itemize}[noitemsep]
  \item Rich progress bars via the \texttt{rich} library.
  \item Consistent \texttt{-o}/\texttt{--output} option with a sensible default.
  \item Clear, coloured error messages on \texttt{stderr}.
  \item Pure-Python dependencies (no external binaries required for
        the new features).
\end{itemize}

% -------------------------------------------------------------------
\section{What Was in Project 1}
% -------------------------------------------------------------------

\subsection{Commands}

\begin{center}
\begin{tabular}{ll}
\toprule
\textbf{Command} & \textbf{Description} \\
\midrule
\texttt{merge}   & Merge two or more PDFs into one file \\
\texttt{split}   & Extract a page range from a PDF \\
\texttt{convert} & Convert images (PNG/JPG/BMP/TIFF) or Office docs to PDF \\
\texttt{info}    & Display metadata (pages, size, author, title, \ldots) \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Dependencies}

\begin{itemize}[noitemsep]
  \item \texttt{pypdf} --- read/write PDF files
  \item \texttt{Pillow} --- image loading and conversion
  \item \texttt{click} --- CLI argument parsing
  \item \texttt{rich} --- terminal output formatting
  \item \texttt{docx2pdf} (optional) --- Office-to-PDF conversion
\end{itemize}

% -------------------------------------------------------------------
\section{What Is New in Project 2}
% -------------------------------------------------------------------

\subsection{New Dependency: reportlab}

The \texttt{watermark} command generates text watermark pages on the
fly.  This requires a PDF canvas library.
\texttt{reportlab}~4.x was added as a new dependency; it is
pure-Python and installs via \texttt{pip}.

\begin{lstlisting}[language=bash]
pip install reportlab
\end{lstlisting}

No other new dependencies were introduced.

\subsection{Module Changes}

\begin{center}
\begin{tabular}{lp{10cm}}
\toprule
\textbf{File} & \textbf{Changes} \\
\midrule
\texttt{pdfuse/\_\_init\_\_.py}  &
  Version bumped from \texttt{0.1.0} to \texttt{0.2.0};
  module docstring updated. \\
\texttt{pdfuse/operations.py}  &
  Added four new functions:
  \texttt{compress\_pdf}, \texttt{rotate\_pdf},
  \texttt{watermark\_pdf} (and helper
  \texttt{\_make\_text\_watermark\_pdf}),
  \texttt{reorder\_pdf}.
  Import of \texttt{Optional} added to the type hints. \\
\texttt{pdfuse/cli.py} &
  Added four new Click commands (\texttt{cmd\_compress},
  \texttt{cmd\_rotate}, \texttt{cmd\_watermark},
  \texttt{cmd\_reorder}) and updated the import block. \\
\bottomrule
\end{tabular}
\end{center}

% -------------------------------------------------------------------
\section{New Commands}
% -------------------------------------------------------------------

\subsection{compress}

\subsubsection*{Purpose}
Reduce file size by applying zlib/deflate compression to every
content stream in the PDF.  This is lossless --- no visual quality is
lost.

\subsubsection*{How it works}
Each page is added to a \texttt{pypdf.PdfWriter} and then
\texttt{compress\_content\_streams()} is called on every page owned
by the writer (the method requires the page to be attached to a
writer before it can be invoked).
Existing metadata is preserved.
The CLI reports the before/after file size and the percentage
reduction.

\subsubsection*{Signature}
\begin{lstlisting}[language=bash]
pdfuse compress FILE [-o OUTPUT]
\end{lstlisting}

\subsubsection*{Options}
\begin{itemize}[noitemsep]
  \item \texttt{FILE} --- input PDF.
  \item \texttt{-o / --output} --- output path (default:
        \texttt{compressed.pdf}).
\end{itemize}

\subsubsection*{Example}
\begin{lstlisting}[language=bash]
pdfuse compress report.pdf -o report_small.pdf
# Output:
# Compressing pages... [############################] 100%
# Compressed report.pdf -> report_small.pdf (2 KB -> 1 KB, 50.0% reduction)
\end{lstlisting}

% -------------------------------------------------------------------
\subsection{rotate}

\subsubsection*{Purpose}
Rotate one or more pages in a PDF by 90, 180, or 270 degrees
(clockwise).  Useful for fixing scanned documents with wrong
orientation.

\subsubsection*{How it works}
\texttt{pypdf} exposes \texttt{page.rotate(angle)} which updates the
\texttt{/Rotate} entry in the page dictionary.  The operation is
applied only to the pages listed with \texttt{--pages}; all other
pages are copied unchanged.

\subsubsection*{Signature}
\begin{lstlisting}[language=bash]
pdfuse rotate FILE --angle {90|180|270} [--pages N[,N...]] [-o OUTPUT]
\end{lstlisting}

\subsubsection*{Options}
\begin{itemize}[noitemsep]
  \item \texttt{--angle} --- rotation in degrees; must be 90, 180, or 270.
  \item \texttt{--pages} --- comma-separated 1-indexed page numbers
        (default: all pages).
  \item \texttt{-o / --output} --- output path (default:
        \texttt{rotated.pdf}).
\end{itemize}

\subsubsection*{Examples}
\begin{lstlisting}[language=bash]
# Rotate all pages 90 degrees clockwise
pdfuse rotate scan.pdf --angle 90 -o scan_fixed.pdf

# Rotate only page 2 by 180 degrees
pdfuse rotate doc.pdf --angle 180 --pages 2 -o doc_p2flipped.pdf

# Rotate pages 1 and 3 by 270 degrees
pdfuse rotate doc.pdf --angle 270 --pages 1,3 -o doc_rotated.pdf
\end{lstlisting}

% -------------------------------------------------------------------
\subsection{watermark}

\subsubsection*{Purpose}
Stamp a watermark on every page of a PDF.  Supports two modes:

\begin{enumerate}[noitemsep]
  \item \textbf{Text watermark} (\texttt{--text}): diagonal,
        semi-transparent grey text centred on each page, generated
        at runtime using \texttt{reportlab}.
  \item \textbf{Stamp watermark} (\texttt{--stamp}): the first page
        of a user-supplied PDF is overlaid on each page (useful for
        logos or custom designs).
\end{enumerate}

\subsubsection*{How it works}
For text watermarks, \texttt{\_make\_text\_watermark\_pdf} renders a
\texttt{reportlab} canvas to an in-memory buffer at the exact
dimensions of each destination page.  The font size is scaled to
10\% of the smaller page dimension (clamped to at least 24~pt).
Text is drawn at a diagonal angle (arctan of height/width) around
the page centre with 30\% opacity.

The watermark page is then merged onto each source page with
\texttt{pypdf}'s \texttt{page.merge\_page()}.  For stamp PDFs the
same merge is applied using the first page of the stamp file.

\subsubsection*{Signature}
\begin{lstlisting}[language=bash]
pdfuse watermark FILE (--text TEXT | --stamp STAMP_PDF) [-o OUTPUT]
\end{lstlisting}

\subsubsection*{Options}
\begin{itemize}[noitemsep]
  \item \texttt{--text TEXT} --- watermark string (e.g.\
        \texttt{CONFIDENTIAL}).
  \item \texttt{--stamp STAMP\_PDF} --- PDF file whose first page is
        used as the stamp.
  \item \texttt{-o / --output} --- output path (default:
        \texttt{watermarked.pdf}).
\end{itemize}

\subsubsection*{Examples}
\begin{lstlisting}[language=bash]
# Add diagonal text watermark
pdfuse watermark contract.pdf --text "DRAFT" -o contract_draft.pdf

# Use a custom PDF stamp (e.g. a logo page)
pdfuse watermark report.pdf --stamp logo.pdf -o report_branded.pdf
\end{lstlisting}

% -------------------------------------------------------------------
\subsection{reorder}

\subsubsection*{Purpose}
Write the pages of a PDF in an arbitrary order specified by the user.
Pages may be repeated, omitted, or rearranged.

\subsubsection*{How it works}
The user provides a comma-separated list of 1-indexed page numbers
via \texttt{--order}.  The function reads each requested page from
the source \texttt{PdfReader} and writes them to a new
\texttt{PdfWriter} in the specified sequence.  Out-of-range page
numbers are caught before writing begins.

\subsubsection*{Signature}
\begin{lstlisting}[language=bash]
pdfuse reorder FILE --order N[,N...] [-o OUTPUT]
\end{lstlisting}

\subsubsection*{Options}
\begin{itemize}[noitemsep]
  \item \texttt{--order} --- comma-separated 1-indexed page list,
        e.g.\ \texttt{3,1,2}.  Pages can be repeated.
  \item \texttt{-o / --output} --- output path (default:
        \texttt{reordered.pdf}).
\end{itemize}

\subsubsection*{Examples}
\begin{lstlisting}[language=bash]
# Reverse a 3-page document
pdfuse reorder doc.pdf --order 3,2,1 -o doc_reversed.pdf

# Move the cover page (page 1) to the end
pdfuse reorder doc.pdf --order 2,3,4,5,1 -o doc_cover_last.pdf

# Duplicate a page (print page 1 twice)
pdfuse reorder doc.pdf --order 1,1,2,3 -o doc_dup_cover.pdf
\end{lstlisting}

% -------------------------------------------------------------------
\section{Summary of Changes}
% -------------------------------------------------------------------

\begin{center}
\begin{tabular}{llp{7.5cm}}
\toprule
\textbf{Version} & \textbf{Commands} & \textbf{Notes} \\
\midrule
0.1.0 & \texttt{merge, split, convert, info} &
  Initial release. \\
0.2.0 & + \texttt{compress, rotate, watermark, reorder} &
  Four new PDF manipulation commands added.
  New dependency: \texttt{reportlab}.
  No breaking changes to existing commands. \\
\bottomrule
\end{tabular}
\end{center}

\end{document}
